---
title: "Installation"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Installation}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

`growthcleanr` has been developed and tested using R version 3.6 and more
recently version 4+. It should work using R on Windows, macOS, or Unix/Linux,
although there are some additional [platform-specific notes](#platform) you may
wish to review.

You can install the `growthcleanr` package directly from GitHub using `devtools`
in the R console with:

```{r, eval = FALSE}
install.packages("devtools")
devtools::install_github("carriedaymont/growthcleanr", ref="main")
```

Note that `ref="main"` is required; the default branch is "main", and must be
referred to explicitly.

If you are unable to install `devtools`, a similar function is available in the
`remotes` package:

```{r, eval = FALSE}
install.packages("remotes")
remotes::install_github("carriedaymont/growthcleanr", ref="main")
```

Note that `ref="main"` is required; the default branch is "main", and must be
referred to explicitly.

`growthcleanr` itself has several dependencies, so it may take a little while to
download and install everything on your machine.

## Optional packages

These packages are not required for running `cleangrowth()` and the other main
functions of `growthcleanr`, but may be necessary for certain use cases.

* `argparser` is used by the `gcdriver.R` script for command line
operation as described in [Working with large data sets](large-data-sets.html)

* `bit64` is necessary if you have long (64-bit) integer `subjid` values

These can be installed the usual way:

```{r, eval = FALSE}
install.packages(c("argparser", "bit64"))
```

## <a name="platform"></a>Platform-specific notes

### Windows

We have observed an issue with using `growthcleanr` with Windows in a large
agency where some R packages are installed by an administrator for a user who
may not have write access permissions on the package folder. Similar issues may
occur with some networked drives. This caused problems with `growthcleanr`'s
parallel processing option. If possible, install R and its packages in locations
hosted on the same local machine and folder(s) for which the primary user has
write permissions. These steps should help to avoid this problem.

Some users have reported that `growthcleanr` runs more slowly on Windows
compared with Linux or macOS.

### macOS

`growthcleanr` uses the
[`data.table`](https://rdatatable.gitlab.io/data.table/) package for R
extensively. `data.table` provides a faster version of R's data frames, and is
used to improve `growthcleanr` performance. Typically `data.table` installs in a
manner that will be able to take advantage of multiple threads. You will know it
worked successfully if, when you load the `data.table` library in R, you see
something like the following:

```{r, eval = FALSE}
library(data.table)
data.table 1.12.2 using 2 threads (see ?getDTthreads).  Latest news: r-datatable.com
```

That data.table reports "using 2 threads" indicates that installation has
succeeded. If the message reports using only one thread, see the advice under
the ["OpenMP enabled compiler for Mac"
instructions](https://github.com/Rdatatable/data.table/wiki/Installation#openmp-enabled-compiler-for-mac)
to re-install `data.table`.

Users have reported errors running multiple batches with `parallel = T` from
within RStudio. If this happens, the problem may be resolved by running from
RGui or from the command line using `Rscript`. An example standalone script that
may be used for this purpose is documented in [Working with large data
sets](large-data-sets.html).

### Docker

This package includes a `Dockerfile` that enables easy installation of R and
`growthcleanr` on a machine with [Docker](https://www.docker.com/) installed. It
requires an up-to-date Docker install, and a few command-line steps, but can save time
over installing R and `growthcleanr`'s dependencies manually.

To install and run `growthcleanr` using Docker, open the PowerShell on Windows, or open
the Terminal on macOS, and enter this `docker` command:

```{bash, eval = FALSE}
% docker run -it mitre/growthcleanr:latest R
```

The first time this command is run, it might take a few minutes to download and extract
several necessary components, but this should be fully automated. If successful, you
should see an R prompt, from which you can use `growthcleanr` as described below.

This R environment is virtualized inside Docker, however, and isolated from your local
machine. Because of this, you will need to map a local folder on your computer into the
Docker environment to work with your own data. For example, if you are on Windows, and
your data is in `C:\Users\exampleuser\analysis`, specify a mapping using the added `-v`
step below:

```{bash, eval = FALSE}
% docker run -it -v C:\Users\exampleusers\analysis:/usr/src/app mitre/growthcleanr:latest R
```

Note that the slashes in file paths reverse direction from the reference to the folder
location on your Windows machine (before the colon) to the folder location on the Docker
container (after the colon); this is intentional, and accounts for how the two different
environments reference disk locations.

Note also that when mapping a folder on Windows, you may be prompted to confirm that you
indeed want to "Share" the folder. This is a standard Windows security practice, and it
is okay to confirm and proceed.

If you are on macOS, and your data is in `/Users/exampleuser/analysis`, specify a folder
mapping like this:

```{bash, eval = FALSE}
% docker run -it -v /Users/exampleuser/analysis:/usr/src/app mitre/growthcleanr:latest R
```

If you mapped a folder, then inside the Docker environment's R prompt, when you then
issue a command like `list.files()`, you should see a list of the same files in the R
session that you see in that folder on your desktop. You can now open and read your data
files, run `cleangrowth()` and other analyses, and write result files to that same
directory.

Exit the Docker R environment with `quit()` as you normally would. Any new files you
saved will appear in the desktop folder you mapped.

## Source-level install for developers

If you want to work with and potentially change the `growthcleanr` code itself,
you can download or clone the `growthcleanr` source code and then install it from
source. To clone the source using `git`:

```{bash, eval = FALSE}
% git clone https://github.com/carriedaymont/growthcleanr.git
```

Either way, once you have the `growthcleanr` package source, open an R session
from the `growthcleanr` base directory. Then install growthcleanr using the R
`devtools` package:

```{r, eval = FALSE}
devtools::install(".")
```

You can also install the package from an installation file if one is obtained.
