---
title: "Installation"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Installation} %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

`growthcleanr` has been developed and tested using R versions 3.6 and 4+. It
should work using R on Windows, macOS, or Unix/Linux, although there are some
additional [platform-specific notes](#platform) you may wish to review.

To get started with `growthcleanr`, install it
[from CRAN](https://cran.r-project.org/package=growthcleanr):

```{r, eval = FALSE}
install.packages("growthcleanr")
```

To install the latest development version from GitHub using `devtools`:

```{r, eval = FALSE}
devtools::install_github("carriedaymont/growthcleanr", ref="main")
```

Installing `growthcleanr` will install several additional packages in turn.

See [GitHub and source-level install for developers](#dev-install) for
additional details.

## Optional packages

These packages are not required for running `cleangrowth()` and the other main
functions of `growthcleanr`, but may be necessary for certain use cases.

- `argparser` is used by the `gcdriver.R` script for command line operation as
  described in [Working with large data sets](large-data-sets.html)

- `bit64` is necessary if you have long (64-bit) integer `subjid` values

These can be installed the usual way:

```{r, eval = FALSE}
install.packages(c("argparser", "bit64"))
```

## <a name="platform"></a>Platform-specific notes

### Windows

We have observed an issue with using `growthcleanr` with Windows in a large
agency where some R packages are installed by an administrator for a user who
may not have write access permissions on the package folder. Similar issues may
occur with some networked drives. This caused problems with `growthcleanr`'s
parallel processing option. If possible, install R and its packages in locations
hosted on the same local machine and folder(s) for which the primary user has
write permissions. These steps should help to avoid this problem.

Some users have reported that `growthcleanr` runs more slowly on Windows
compared with Linux or macOS.

### macOS

`growthcleanr` uses the [`data.table`](https://rdatatable.gitlab.io/data.table/)
package for R extensively. `data.table` provides a faster version of R's data
frames, and is used to improve `growthcleanr` performance. Typically
`data.table` installs in a manner that will be able to take advantage of
multiple threads. You will know it worked successfully if, when you load the
`data.table` library in R, you see something like the following:

```{r, eval = FALSE}
library(data.table)
data.table 1.12.2 using 2 threads (see ?getDTthreads).  Latest news: r-datatable.com
```

That data.table reports "using 2 threads" indicates that installation has
succeeded. If the message reports using only one thread, see the advice under
the
["OpenMP enabled compiler for Mac" instructions](https://github.com/Rdatatable/data.table/wiki/Installation#openmp-enabled-compiler-for-mac)
to re-install `data.table`.

Users have reported errors running multiple batches with `parallel = T` from
within RStudio. If this happens, the problem may be resolved by running from
RGui or from the command line using `Rscript`. An example standalone script that
may be used for this purpose is documented in
[Working with large data sets](large-data-sets.html).

### Docker

This package includes a `Dockerfile` that enables easy installation of R and
`growthcleanr` on a machine with [Docker](https://www.docker.com/) installed. It
requires an up-to-date Docker install, and a few command-line steps, but can
save time over installing R and `growthcleanr`'s dependencies manually.

To install and run `growthcleanr` using Docker, open the PowerShell on Windows,
or open the Terminal on macOS, and enter this `docker` command:

```{bash, eval = FALSE}
% docker run -it ghcr.io/mitre/growthcleanr/gcr-image:latest R
```

The image tag `latest` in the example above will refer to the latest version of
the package available on the main branch of the
[mitre/growthcleanr](https://github.com/mitre/growthcleanr) repository, which is
typically in close sync with the upstream
[carriedaymont/growthcleanr](https://github.com/carriedaymont/growthcleanr)
repository. This will usually be the latest released version of `growthcleanr`.
To explicitly choose a release by name, replace `latest` with the release tag,
e.g. for the released package `v2.1.0`:

```{bash, eval = FALSE}
% docker run -it ghcr.io/mitre/growthcleanr/gcr-image:v2.1.0 R
```

Whichever package you use, the first time this command is run, it might take a
few minutes to download and extract several necessary components, but this
should be fully automated. If successful, you should see an R prompt, from which
you can use the `growthcleanr` package immediately.

This R environment is virtualized inside Docker, however, and isolated from your
local machine. Because of this, you will need to map a local folder on your
computer into the Docker environment to work with your own data. For example, if
you are on Windows, and your data is in `C:\Users\exampleuser\analysis`, specify
a mapping using the added `-v` step below:

```{bash, eval = FALSE}
% docker run -it -v C:\Users\exampleusers\analysis:/usr/src/app \
    ghcr.io/mitre/growthcleanr/gcr-image:latest R
```

Note that the slashes in file paths reverse direction from the reference to the
folder location on your Windows machine (before the colon) to the folder
location on the Docker container (after the colon); this is intentional, and
accounts for how the two different environments reference disk locations.

Note also that when mapping a folder on Windows, you may be prompted to confirm
that you indeed want to "Share" the folder. This is a standard Windows security
practice, and it is okay to confirm and proceed.

If you are on macOS, and your data is in `/Users/exampleuser/analysis`, specify
a folder mapping like this:

```{bash, eval = FALSE}
% docker run -it -v /Users/exampleuser/analysis:/usr/src/app \
    ghcr.io/mitre/growthcleanr/gcr-image:latest R
```

If you mapped a folder, then inside the Docker environment's R prompt, when you
then issue a command like `list.files()`, you should see a list of the same
files in the R session that you see in that folder on your desktop. You can now
open and read your data files, run `cleangrowth()` and other analyses, and write
result files to that same directory.

Exit the Docker R environment with `quit()` as you normally would. Any new files
you saved will appear in the desktop folder you mapped.

## <a name="dev-install"></a>GitHub and source-level install for developers

You can install the `growthcleanr` package directly from GitHub using `devtools`
in the R console with:

```{r, eval = FALSE}
install.packages("devtools")
devtools::install_github("carriedaymont/growthcleanr", ref="main")
```

`growthcleanr` itself has several dependencies, so it may take a little while to
download and install everything on your machine.

Note that the `ref="main"` part is required; the default value of `ref` refers
to a branch name that is not used in the `growthcleanr` repository, which
instead uses a default branch called "`main`".

To install a different branch, for example if you want to test a branch
associated with a merge request, specify the branch name as the value of `ref`.

If you are unable to install `devtools`, a similar function is available in the
`remotes` package:

```{r, eval = FALSE}
install.packages("remotes")
remotes::install_github("carriedaymont/growthcleanr", ref="main")
```

If you are developing the `growthcleanr` code itself, you can download or clone
the `growthcleanr` source code and then install it from source. To clone the
source using `git`:

```{bash, eval = FALSE}
% git clone https://github.com/carriedaymont/growthcleanr.git
```

Once you have the `growthcleanr` package source, open an R session from the
`growthcleanr` base directory. Then install growthcleanr using the R `devtools`
package:

```{r, eval = FALSE}
devtools::install(".")
```
