---
title: "Next steps"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Next steps}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

While both the pediatric and adult algorithms are considered complete, we have
identified several areas for additional research and potential enhancement.

# Pediatric algorithm - carried forward adjustments

An experiment in re-assessing pediatric height measurements initially marked as
`Exclude-Carried-Forward` based on growth velocity is included in the current
version of this package.

Because of the complexity involved in this approach, this is implemented as an
independent function with its own driver script for further study. Our hope is
to revise this strategy, and, if feasible, incorporate it into the main
`cleangrowth()` algorithm.

## Running the experiment

The primary function is called `adjustcarryforward()`, and is implemented in the
file `R/adjustcarryforward.R`. It is separate from the main `R/growth.R` file as
it is set up to be run separately, using the output from `cleangrowth()`. The
`adjustcarryforward()` function is available in the main package namespace.

The primary tools for running this function are the script `exec/testadjustcf.R` and the function `testacf()`. 
These will run the experimental function on an existing dataset (specified as a
CSV file or data frame) and pass it a sweep of parameter values for testing the new strategy.
This tool will produce a combined output file or list that includes the original result
alongside the resulting re-inclusion determination for each measurement for each
parameter value run.

For example, if you have run `cleangrowth()` on the `syngrowth` synthetic data
set provided with `growthcleanr` as described in the main `README.md` file, save
it to a CSV file for `exec/testadjustcf.R`:

```R
> fwrite(cleaned_data, "cleaned.csv", row.names = F)
```
For `testacf()`, keep cleaned_data in your environment. Note that the column names should be as described for `cleaned_data` the Example under Quickstart.

To use `exec/testadjustcf.R`, first navigate to the growthcleanr package directory in the command line. Then
execute the sweep script is executed from the command line on the cleaned data
file for as follows (with the assumption that cleaned.csv is in your current
directory; otherwise, make sure to specify the path relative to your current
directory):

```bash
% Rscript exec/testadjustcf.R cleaned.csv
```

To use `testacf()`, run the following in the console (with `cleaned_data` in the environment):

```{r, eval=FALSE}
result_list <- testacf(cleaned_data)
```

This `result_list` will contain 2 or 3 objects;
- `testacf_res`: a data frame with adjustcarryforward results for each run
- `params`: a data frame containing the parameters for each run
- `debug_filtered_data`: data frame with original data, returned if debug is `TRUE`

By default, the script/function will generate a range of values with nine steps for the
following parameters, where the min and max surround the default value:

| parameter          | default | min | max  |
| ------------------ | ------- | --- | ---- |
| `minfactor`        | 0.5     | 0   | 1    |
| `maxfactor`        | 2       | 0   | 4    |
| `banddiff`         | 3       | 0   | 6    |
| `banddiff_plus`    | 5.5     | 0   | 11   |
| `min_ht.exp_under` | 2       | 0   | 4    |
| `min_ht.exp_over`  | 0       | -1  | 1    |
| `max_ht.exp_under` | 0.33    | 0   | 0.66 |
| `max_ht.exp_over`  | 1.5     | 0   | 3    |

The determination of these values depends on the search type (specified with the
option `--searchtype` (script)/argument `searchtype` (function):

- `random` (default): Values will be generated randomly, with equal amounts of values
  on either side of the midpoint. The midpoint is always included.
  - Note that if an even number is specified for `--gridlength`/`grid.length`, one will be added to
    include the midpoint in the run.
  - A random seed can be specified with `--seed`/`seed` (default 7).
- `line-grid`: Values will be evenly distributed along the range for each parameter.
  If the `--gridlength`/`grid.length` specified is odd, this will include the midpoint.
- `full-grid`: Values for each included parameter will evenly distributed along the
  range for each parameter and in a full combination between all parameters.

  - Thus, the amount of runs done will be the `--gridlength`/`gridlength`^(number of included
    parameters).
  - Default includes a full grid search among all parameters. To specify use of only
    specific parameters, use the `--param`/`param` option, which specifies a CSV of the following
    format:

  | parameter          | include | value |
  | ------------------ | ------- | ----- |
  | `minfactor`        | T       |       |
  | `maxfactor`        | F       | 3     |
  | `banddiff`         | F       |       |
  | `banddiff_plus`    | F       |       |
  | `min_ht.exp_under` | T       |       |
  | `min_ht.exp_over`  | F       |       |
  | `max_ht.exp_under` | F       | .5    |
  | `max_ht.exp_over`  | F       |       |

  - The first column specifies all the parameter names; the second specifies a
    true or false value for whether or not the parameter should be included; the
    third specifies a constant value to be used for not included parameters, left
    empty if the value should be the default.
  - In the above example, `minfactor` and `min_ht.exp_under` will be included, and
    `maxfactor` and `max_ht_exp_under` will not be included, but will use 3 and .5
    as their values.
  - Warning: this will take much longer!

The default number of sweep steps is 9; this can be changed with the option
`--gridlength`/`gridlength`.

For testing options of handling strings of multiple carried forward values,
several options from 0 to 3 have been incorporated. 0 (no change) is the default
option, and can be changed `--exclude_opt`/`exclude_opt`. More information on each option can
be found in the `adjustcarryforward()` documentation.

In addition to multiple options for carried-forward strings, "answers" for a
given dataset have been incorporated. When the `--add_answers` flag/`add_answers` argument is set to
`TRUE` (`TRUE` by default), a column called `acf_answers` will have, for each
height value, "Definitely Exclude", "Definitely Include", or "Unknown" (if it
does not fall in either category). Weight values are set as `NA`.

For example, for a 9-step sweep with the default search type, `random`, the
parameters passed to the function in each pass will be:

```R
run	minfactor	maxfactor	banddiff	banddiff_plus	min_ht.exp_under	min_ht.exp_over	max_ht.exp_under	max_ht.exp_over
1	0.494454649	0.331710969	1.681997601	5.438065292	0.371428523	-0.200524185	0.296497153	0.244186167
2	0.198872727	0.918207332	0.0261138	0.361051567	0.370286443	-0.618056939	0.318943811	0.280842425
3	0.057848889	0.343496154	2.957211272	3.448713172	0.758593493	-0.240298769	0.189112731	0.586874211
4	0.034874339	0.462954204	0.949754412	2.697612725	1.694048784	-0.563224398	0.237626234	0.410851815
5	0.5	2	3	5.5	2	0	0.33	1.5
6	0.621874695	3.545623892	4.918346826	10.84063427	2.996152267	0.904217721	0.585439346	1.787876622
7	0.896005213	2.192603083	3.885669706	7.492214661	3.581171147	0.319534914	0.537161064	2.25658771
8	0.670031176	2.90689554	5.990111082	9.239964036	3.676927744	0.082569093	0.568586483	2.645760535
9	0.98603125	2.169401426	5.718063961	6.950459617	2.91380773	0.816289079	0.457654322	2.540503306
```

In a 9-step sweep with a `line-grid` search type, the parameters passed to the
function in each pass will be:

```R
run  minfactor  maxfactor  banddiff  banddiff_plus  min_ht.exp_under  min_ht.exp_over  max_ht.exp_under  max_ht.exp_over
1    0          0          0         0              0                 -1               0                 0
2    0.125      0.5        0.75      1.375          0.5               -0.75            0.0825            0.375
3    0.25       1          1.5       2.75           1                 -0.5             0.165             0.75
4    0.375      1.5        2.25      4.125          1.5               -0.25            0.2475            1.125
5    0.5        2          3         5.5            2                 0                0.33              1.5
6    0.625      2.5        3.75      6.875          2.5               0.25             0.4125            1.875
7    0.75       3          4.5       8.25           3                 0.5              0.495             2.25
8    0.875      3.5        5.25      9.625          3.5               0.75             0.5775            2.625
9    1          4          6         11             4                 1                0.66              3
```

In a 3-step sweep with a `full-grid` search type, with the `--param` CSV/`param` data frame specified as
in the above example, the parameters passed to the function in each pass will be:

```R
run minfactor maxfactor banddiff banddiff_plus min_ht.exp_under min_ht.exp_over max_ht.exp_under max_ht.exp_over
1       0.0         3        3           5.5                0               0              0.5             1.5
2       0.5         3        3           5.5                0               0              0.5             1.5
3       1.0         3        3           5.5                0               0              0.5             1.5
4       0.0         3        3           5.5                2               0              0.5             1.5
5       0.5         3        3           5.5                2               0              0.5             1.5
6       1.0         3        3           5.5                2               0              0.5             1.5
7       0.0         3        3           5.5                4               0              0.5             1.5
8       0.5         3        3           5.5                4               0              0.5             1.5
9       1.0         3        3           5.5                4               0              0.5             1.5
```

For the script, the output in the working directory will contain the sweep parameters, like the above,
in a file called `test_adjustcarrforward_DATE_TIME_parameters.csv`, and the output with
adjustment results in a file called `test_adjustcarrforward_DATE_TIME.csv`, where DATE
and TIME are the system date and time. For the function, a list will be returned with the parameters as a data.frame in the `params` entry and output adjustment results as a data.frame in the `testacf_res` entry.

For example, a 5-step sweep with the `line-grid` search would be run with this command:

```bash
% Rscript exec/textadjustcf.R --gridlength 5 --searchtype line-grid cleaned.csv
```

or with this function execution:

```{r, eval=FALSE}
result_list <- testacf(
  cleaned_data,
  grid.length = 5,
  searchtype = "line-grid"
)
```

The parameter set for the sweep in file `test_adjustcarrforward_DATE_TIME_parameters.csv` (script)/`params` data frame of `result_list` (function) would be:

```R
run  minfactor  maxfactor  banddiff  banddiff_plus  min_ht.exp_under  min_ht.exp_over  max_ht.exp_under  max_ht.exp_over
1    0          0          0         0              0                 -1               0                 0
2    0.25       1          1.5       2.75           1                 -0.5             0.165             0.75
3    0.5        2          3         5.5            2                 0                0.33              1.5
4    0.75       3          4.5       8.25           3                 0.5              0.495             2.25
5    1          4          6         11             4                 1                0.66              3
```

Note that an odd-numbered length will include the default values in the middle run of the
sweep (hence the examples with 5 and 9 step sweeps).

And the first few result rows in `test_adjustcarrforward_DATE_TIME.csv` (script)/`testacf_res` data frame of `result_list` (function) would be:

```R
id     subjid    sex  agedays  param     measurement  gcr_result                      run-1      run-2      run-3      run-4      run-5
1510   775155    0    889      HEIGHTCM  84.9         Exclude-Extraneous-Same-Day Missing    Missing    Missing    Missing    Missing
1511   775155    0    889      HEIGHTCM  89.06        Include                     No Change  No Change  No Change  No Change  No Change
1512   775155    0    1071     HEIGHTCM  92.5         Include                     No Change  No Change  No Change  No Change  No Change
1513   775155    0    1253     HEIGHTCM  96.2         Include                     No Change  No Change  No Change  No Change  No Change
1514   775155    0    1435     HEIGHTCM  96.2         Exclude-Carried-Forward     No Change  No Change  Include    Include    Include
1515   775155    0    1435     HEIGHTCM  99.692       Include                     No Change  No Change  No Change  No Change  No Change
1516   775155    0    1806     HEIGHTCM  106.1        Include                     No Change  No Change  No Change  No Change  No Change
1517   775155    0    2177     HEIGHTCM  112.3        Include                     No Change  No Change  No Change  No Change  No Change
1518   775155    0    889      WEIGHTKG  13.1         Include                     No Change  No Change  No Change  No Change  No Change
```

The fifth row in the example above demonstrates the results of the experimental
script; for runs 1 and 2, the result is not changed, but for runs 3-5, the
measurement is adjusted for reinclusion. To demonstrate the range, the following
is an extract of measurements only marked as carried forward exclusions by
`cleangrowth()`:

```R
id     subjid     sex  agedays  param     measurement  gcr_result                   run-1      run-2      run-3      run-4      run-5
1514   775155     0    1435     HEIGHTCM  96.2         Exclude-Carried-Forward  No Change  No Change  Include    Include    Include
1521   775155     0    1435     WEIGHTKG  15.3         Exclude-Carried-Forward  No Change  No Change  No Change  No Change  No Change
7952   1340377    1    1806     HEIGHTCM  107.1        Exclude-Carried-Forward  No Change  Include    Include    Include    Include
7967   1340377    1    1806     WEIGHTKG  18.4         Exclude-Carried-Forward  No Change  No Change  No Change  No Change  No Change
41775  3643526    1    1253     HEIGHTCM  87.808       Exclude-Carried-Forward  Include    Include    Include    Include    Include
44901  3706097    0    4032     HEIGHTCM  138.8        Exclude-Carried-Forward  No Change  Include    Include    Include    Include
30011  5792371    1    3661     HEIGHTCM  145.4        Exclude-Carried-Forward  No Change  Include    Include    Include    Include
30013  5792371    1    4032     HEIGHTCM  145.4        Exclude-Carried-Forward  No Change  No Change  No Change  No Change  No Change
30016  5792371    1    1071     WEIGHTKG  15.9         Exclude-Carried-Forward  No Change  No Change  No Change  No Change  No Change
```

Some of these values are not adjusted at all; one is from run 1 on, a few are
from run 2 on, and one is from run 3 on.

# Adult algorithm - potential enhancements

The [adult algorithm](adult-algorithm.html) added in release 2.0 is complete, but
has some known limitations that present areas for future research and
enhancement. These areas are enumerated below.

We welcome your ideas on these or any additional enhancements via email,
tickets, or pull requests.

Any active work on these areas will be ticketed and managed through GitHub.

## "Evil twins"

In some cases, runs of two or three similar - but not necessarily identical -
deviant weights can occur, which are difficult to detect. One approach to
address these could involve modifying the moderate EWMA step, but capturing
these reliably may be a very challenging task.

## Moderate EWMA adjustments

A number of adjustments are possible to improve this step, such as tweaks to the
weight change allowance, polation, and prioritization. This might require fewer
changes to code but more extensive testing to ensure confidence in new threshold
numbers/details.

## Loosen criteria on hundreds, unit errors, etc.

The criteria for some of these steps may be loosened to catch more implausible
values. For unit error correction, we could consider having two levels: one
suitable for correction, the other not.

## Number of height loss events

While maintaining overall height loss limits, we could allow for the possibility
of more height loss events rather than assuming height loss will occur in no
more than two big jumps.

## Error load

We could consider a change so that error load is not triggered if the only
exclusions are for heights 3+D outside of w2.
