---
title: "Quickstart"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Quickstart}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

## R setup

To get started with `growthcleanr`, the `growthcleanr` package must be
installed. Installing `growthcleanr` will in turn install several additional
packages. `growthcleanr` is not currently on CRAN.

One option for installation is using the `devtools` package. Install `devtools`
and `growthcleanr` with the following commands:

```{r, eval = FALSE}
install.packages("devtools")
devtools::install_github("carriedaymont/growthcleanr", ref="main")
```

If you are unable to install `devtools`, a similar function is available in the
`remotes` package:

```{r, eval = FALSE}
install.packages("remotes")
remotes::install_github("carriedaymont/growthcleanr", ref="main")
```

Note that `ref="main"` is required whether you use `devtools` or `remotes`. The
default `growthcleanr` code branch is "main", and this must be referred to
explicitly.

Further installation details and notes can be found under
[Installation](installation.html).

## Data preparation

Once the `growthcleanr` package and its dependencies are installed, the minimal
input data vectors needed for cleaning data are:

1. `subjid` - a unique identifier for each subject; if you have long (64-bit)
   integer `subjid` values, add the `bit64` R package to your install list to
   ensure R will process these correctly
2. `param` - a string designating the measurement being '`HEIGHTCM`',
   '`LENGTHCM`', or '`WEIGHTKG`'
3. `agedays` - a numeric value for age in days at time of measurement; will be
   rounded down (floor) to nearest integer if needed (note that there are
   30.4375 days in one month; see also the additional
   [notes on calculating `agemos`](https://www.cdc.gov/nccdphp/dnpao/growthcharts/resources/sas.htm)
   from whole numbers of months
4. `sex` - a numeric value of `0` (male) or `1` (female)
5. `measurement` - numeric value of measurement in cm for height or kg for
   weight

**Note**: the dataset should be sorted by `subjid`, `param`, and `agedays`, as
demonstrated in the Example below.

## Example

Starting with a dataset such as the following:

| subjid | param    | agedays | sex | measurement |
| ------ | -------- | ------- | --- | ----------- |
| 1      | HEIGHTCM | 2790    | 0   | 118.5       |
| 1      | HEIGHTCM | 3677    | 0   | 148.22      |
| 1      | WEIGHTKG | 2790    | 0   | 23.8        |
| 1      | WEIGHTKG | 3677    | 0   | 38.41       |
| 2      | HEIGHTCM | 2112    | 1   | 118.2       |
| 2      | HEIGHTCM | 2410    | 1   | 123.9       |
| 2      | HEIGHTCM | 2708    | 1   | 128.03      |
| 2      | HEIGHTCM | 3029    | 1   | 135.3       |
| 2      | WEIGHTKG | 2112    | 1   | 25.84       |
| 2      | WEIGHTKG | 2410    | 1   | 28.61       |
| 2      | WEIGHTKG | 2708    | 1   | 28.61       |
| 2      | WEIGHTKG | 3029    | 1   | 34.5        |
| ...    | ...      | ...     | ... | ...         |

Note that in this example, we see data from two subjects, sorted as described
above, with two pairs of height/weight measurements for the first patient and
four pairs for the second. Most of these values pass the "eyeball test", but the
third weight measurement for `subjid` 2 looks like it might be a value that has
been carried forward from the previous value, given that their same ageday
height has changed by several centimeters but their third height value is
exactly the same as their second. This is an example of an implausible value and
exclusion type that `growthcleanr` might identify.

The `cleangrowth()` function provided by `growthcleanr` will return a vector
that specifies, for each measurement, whether to:

- "`Include`" the measurement,
- Conclude that the measurement is "`Missing`",
- Or to "`Exclude`" the measurement (numerous variations of `Exclude` exist, as
  specified in [Understanding growthcleanr output](output.html)).

For a data.frame object `source_data` containing growth data:

```{r, eval = FALSE}
library(growthcleanr)

# prepare data as a data.table
data <- as.data.table(source_data)

# set the data.table key for better indexing
setkey(data, subjid, param, agedays)

# generate new exclusion flag field using function
cleaned_data <- data[, gcr_result := cleangrowth(subjid, param, agedays, sex, measurement)]

# extract data limited only to values flagged for inclusion:
only_included_data <- cleaned_data[gcr_result == "Include"]
```

If our example dataset above were named `source_data`, examining `cleaned_data`
would show:

| subjid | param    | agedays | sex | measurement | gcr_result              |
| ------ | -------- | ------- | --- | ----------- | ----------------------- |
| 1      | HEIGHTCM | 2790    | 0   | 118.5       | Include                 |
| 1      | HEIGHTCM | 3677    | 0   | 148.22      | Include                 |
| 1      | WEIGHTKG | 2790    | 0   | 23.8        | Include                 |
| 1      | WEIGHTKG | 3677    | 0   | 38.41       | Include                 |
| 2      | HEIGHTCM | 2112    | 1   | 118.2       | Include                 |
| 2      | HEIGHTCM | 2410    | 1   | 123.9       | Include                 |
| 2      | HEIGHTCM | 2708    | 1   | 128.03      | Include                 |
| 2      | HEIGHTCM | 3029    | 1   | 135.3       | Include                 |
| 2      | WEIGHTKG | 2112    | 1   | 25.84       | Include                 |
| 2      | WEIGHTKG | 2410    | 1   | 28.61       | Include                 |
| 2      | WEIGHTKG | 2708    | 1   | 28.61       | Exclude-Carried-Forward |
| 2      | WEIGHTKG | 3029    | 1   | 34.5        | Include                 |
| ...    | ...      | ...     | ... | ...         | ...                     |

In this sample output, we see that `growthcleanr` has indeed flagged the third
weight measurement for `subjid 2` as an apparent carried forward value. The
measurement remains in the dataset, and it is up to the researcher to determine
which exclusions to accept directly or investigate further.

This example uses the default configuration options for `cleangrowth()`. Longer
examples demonstrating some common options are in [Usage](usage.html), and full
details about all options are documented in
[Configuration](configuration.html).

To compute percentiles and Z-scores using the included CDC macro, continue with
these next steps:

```{r, eval = FALSE}
# Use the built-in utility function to convert the input observations to wide
# format for BMI calculation
cleaned_data_wide <- longwide(cleaned_data)

# Compute simple BMI
cleaned_data_bmi <- simple_bmi(cleaned_data_wide)

# Compute Z-scores and percentiles
cleaned_data_bmiz <- ext_bmiz(cleaned_data_bmi)
```

The `longwide()` function defaults to converting only records included by
`cleangrowth()`, with options to change this default, and `ext_bmiz()` will
return the wide version of the cleaned data with many metrics added. Both of
these functions are described in
[Computing BMI percentiles and Z-scores](utilities.html).