---
title: "Usage"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

With the `growthcleanr` package installed as described in
[Installation](installation.html), it can be loaded as a package:

```{r}
library(growthcleanr)
```

## Basic operations using example synthetic data

For convenience, `growthcleanr` ships with an example synthetic dataset created
using [Synthea](https://synthetichealth.github.io/synthea/), with support for
simulated growth measurement errors based on the protocol included as
supplementary material to the [Daymont et al. paper](#cite). This dataset is
automatically loaded with the `growthcleanr` library, and is called `syngrowth`.

```{r}
dim(syngrowth)
head(syngrowth)
```

It can be used as in the following example. Note that processing the example data
with `cleangrowth()` will likely take a few minutes to complete.

```{r, eval = FALSE}
library(data.table)
library(dplyr)

# Convert the `syngrowth` data frame to a `data.table`
data <- as.data.table(syngrowth)

# `setkey()` creates an efficient sorting key on the `data.table`; this is required
# for `cleangrowth()`
setkey(data, subjid, param, agedays)

# Add a column `gcr_result` using `cleangrowth`
cleaned_data <- data[, gcr_result := cleangrowth(subjid, param, agedays, sex, measurement)]

# View a sample of the results
head(cleaned_data)
      id                               subjid sex  agedays    param measurement                  gcr_result
1: 83330 002986c5-354d-bb9d-c180-4ce26813ca28   1 20489.22 HEIGHTCM       151.1                     Include
2: 83332 002986c5-354d-bb9d-c180-4ce26813ca28   1 20860.22 HEIGHTCM       151.1                     Include
3: 83334 002986c5-354d-bb9d-c180-4ce26813ca28   1 20860.22 HEIGHTCM       150.6 Exclude-Same-Day-Extraneous
4: 83335 002986c5-354d-bb9d-c180-4ce26813ca28   1 21231.22 HEIGHTCM       151.1                     Include
5: 83337 002986c5-354d-bb9d-c180-4ce26813ca28   1 21602.22 HEIGHTCM       151.1                     Include
6: 83339 002986c5-354d-bb9d-c180-4ce26813ca28   1 21623.22 HEIGHTCM       151.1                     Include

# Summarize results by result type
cleaned_data %>% group_by(gcr_result) %>% tally(sort=TRUE)
# A tibble: 26 x 2
   gcr_result                      n
   <fct>                       <int>
 1 Include                     61652
 2 Exclude-Extraneous-Same-Day 11263
 3 Exclude-Carried-Forward      7093
 4 Exclude-Same-Day-Extraneous  4010
 5 Exclude-Same-Day-Identical    623
 6 Exclude-SD-Cutoff             175
 7 Exclude-EWMA-8                139
 8 Exclude-Distinct-3-Or-More    125
 9 Exclude-BIV                   108
10 Exclude-EWMA-Extreme           99
# … with 16 more rows
```

If you are able to run these steps and see a similar result, you have the
`growthcleanr` package installed correctly. The resulting `cleaned_data` can be
reviewed, subsetted, and compared in more detail using all the tools R provides.

## Basic configuration options

For complete information about the options that can be set on the
`cleangrowth()` function, see [Configuration](#configuration). Below are a few
additional examples:

This example shows three configuration options in use:

```{r, eval = FALSE}
cleaned_data <- data[,
  gcr_result_both := cleangrowth(
    subjid, param, agedays, sex, measurement,
    lt3.exclude.mode = "flag.both",
    ref.data.path = "inst/extdata/",
    quietly = FALSE
  )
]
```

- `lt3.exclude.mode = "flag.both"` will exclude both measurements for a
  subject if they only have two unexcluded measurements of one parameter type
  with at least one implausible value and no same age-day measurements of the
  other parameter.

- `ref.data.path = "inst/extdata"` shouldn't be necessary if you have the
  `growthcleanr` package installed, but if you are running it from its
  source directly you may need to specify its full path.

- `quietly = F` enables verbose output, marking the progress of the algorithm
  through its many processing steps. This can be very helpful while testing.

This example shows built-in options for processing data in parallel batches,
which can speed the process while working with large data sets:

```{r, eval = FALSE}
cleaned_data <- data[,
  gcr_result_both := cleangrowth(
    subjid, param, agedays, sex, measurement,
    parallel = TRUE,
    num.batches = 4,
    log.path = "logs"
  )
]
```

- `parallel` tells `cleangrowth()` to run in parallel; the default is `FALSE`

- `num.batches` specifies how many batches to process in parallel

The best `num.batches` value for your environment may vary depending on the
computing resources you have available. If you do not specify `num.batches`,
`growthcleanr` will estimate a batch count based on R functions for checking the
system hardware. If you are working with large datasets, you may need to
experiment with these options to determine the best settings for your needs. You
may also find it helpful to review the additional notes on
[Working with large datasets](large-data-sets.html).

The default value of `log.path` is `"."`, the current working directory.
`growthcleanr` will write batch-specific log files to the `log.path` directory,
and will create the directory if necessary.

Note that if you run `growthcleanr` in parallel with multiple batches and see
warning errors such as the following, they can be ignored.

```{r, eval = FALSE}
Warning messages:
1: <anonymous>: ... may be used in an incorrect context: ‘.fun(piece, ...)’
2: <anonymous>: ... may be used in an incorrect context: ‘.fun(piece, ...)’
```

This set of warnings is coming from one of `growthcleanr`'s dependencies, and
does not indicate either failure or improper execution.

This final example demonstrates using two option specific to adult subjects,
`adult_cutpoint` and `weight_cap`.

```{r, eval = FALSE}
cleaned_data <- data[,
  gcr_result_both := cleangrowth(
    subjid, param, agedays, sex, measurement,
    adult_cutpoint = 18,
    weight_cap = 181.4
  )
]
```

- `adult_cutpoint` defines a subject age in years at which to cut `growthcleanr`
  analysis into separate sets for pediatric and adult subjects. The default is
  20 years, and this may be lowered to 18, as shown above.

- `weight_cap` specifies a known hard limit clamp for weight values. sometimes
  data shared for research may have measurement values clamped to hard limits
  for privacy protection. The growthcleanr adult algorithm can be configured to
  watch for a known weight cap and adjust its assessment accordingly.

The full algorithm for assessing adult measurements is described in detail in
[Adult algorithm](adult-algorithm.html).