---
title: "Utilities for computing pediatric BMI percentiles, Z-scores, and related tools"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Utilities}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---


The CDC provides a
[SAS macro for computing BMI percentiles and Z-scores](https://www.cdc.gov/nccdphp/dnpao/growthcharts/resources/sas.htm).
The function `ext_bmiz()`, included in `growthcleanr`, provides an equivalent
feature. `ext_bmiz()` calculates the sigma (scale parameter for the half-normal
distribution, extended BMI percentile), extended BMIz, and the CDC LMS z-scores
for weight, height, and BMI. Note that for BMIs ≤ 95th percentile of the CDC
growth charts, the extended values for BMI are equal to the LMS values. The
extended values differ only for children who have a BMI > 95th percentile.

The function assumes a variable ‘sex’ (coded as 1=boys / 2=girls) and variables for age
in months, weight (kg), height (cm), and BMI (weight/ht2). Please be careful with age -
the units should be months and use the most accurate information available (e.g.,
23.4928 months. The extended BMIz is the inverse cumulative distribution function (CDF)
of the extended BMI percentile. If the extended percentile is very close to 100, the
`qnorm` function in R produces an infinite value. The occurs only if the extended BMI
percentile is > 99.99999999999999. This occurs infrequently, such as a 48-month-old with
a BMI > 39, and it is likely that these BMIs represent data entry errors. For these
cases, extended BMIz is set to 8.21, a value that is slightly greater than the largest
value that can be calculated.

The CDC SAS macro was updated in December 2022, and `ext_bmiz()` has been
updated to match it as of growthcleanr v2.1.0.

## Converting long growthcleanr data to wide format w/BMI

Because `ext_bmiz()` performs cross-sectional analysis of BMI, observation data must be
in a wide format, i.e. with height and weight information on the same row. This is
distinct from `cleangrowth()`, which performs longitudinal analysis on all observations
for each subject, presented in a long format with one observation per row. To facilitate
use of both functions, `growthcleanr` includes utility functions to transform data used
with `cleangrowth()` for use with `ext_bmiz()`. They are optimized to move data directly
from the output of `cleangrowth()` into input for `ext_bmiz()`, but have options to
support independent use as well.

Using the `syngrowth` example dataset, to convert the data after it has been cleaned by
`cleangrowth()` for use with `ext_bmiz()`, use `longwide()` and `simple_bmi()`:

```{r, eval = FALSE}
# Use the built-in utility function to convert the input observations to wide
# format for BMI calculation
cleaned_data_wide <- longwide(cleaned_data)

# Compute simple BMI values (adds column "bmi")
cleaned_data_bmi <- simple_bmi(cleaned_data_wide)

# Compute Z-scores and percentiles
cleaned_data_bmiz <- ext_bmiz(cleaned_data_bmi)
```

Note that this assumes that `cleaned_data` has the same structure as described
in [Quickstart - Data preparation](quickstart.html#data-preparation-1):

```{r, eval = FALSE}
names(cleaned_data)
[1] "id"          "subjid"      "sex"         "agedays"     "param"       "measurement" "gcr_result"
```

The wide dataset `cleaned_data_wide` will include rows with aligned height and weight
measurements drawn from the observations in `cleaned_data` marked by `cleangrowth()` for
inclusion. As such, it will be a shorter dataset (fewer rows) based on fewer
observations.

```{r, eval = FALSE}
dim(cleaned_data)
[1] 85728     7

dim(cleaned_data_wide)
[1] 26701     9

head(cleaned_data_wide)
                                subjid    agey     agem sex   wt wt_id    ht ht_id  agedays
1 002986c5-354d-bb9d-c180-4ce26813ca28 56.0964 673.1568   2 71.7 83331 151.1 83330 20489.22
2 002986c5-354d-bb9d-c180-4ce26813ca28 57.1122 685.3464   2 73.2 83333 151.1 83332 20860.22
3 002986c5-354d-bb9d-c180-4ce26813ca28 58.1279 697.5348   2 74.6 83336 151.1 83335 21231.22
4 002986c5-354d-bb9d-c180-4ce26813ca28 59.1437 709.7244   2 72.8 83338 151.1 83337 21602.22
5 002986c5-354d-bb9d-c180-4ce26813ca28 59.2012 710.4144   2 72.4 83340 151.1 83339 21623.22
6 002986c5-354d-bb9d-c180-4ce26813ca28 60.1594 721.9128   2 69.4 83343 151.1 83342 21973.22
```

In this example, the subject identifiers previously marked as `subjid` are now in the
`id` column; individual identifiers for observations of a single parameter are not
present.

`longwide()` can be called with name mapping parameters if your input set uses different
column names. For example, if `my_cleaned_data` specifies age in days as `aged` and
parameter type as `type`, specify each, with quotes:

```{r, eval = FALSE}
head(my_cleaned_data)
     id subjid sex    aged     type measurement                  gcr_result
1: 1510 775155   0     889 HEIGHTCM       84.90 Exclude-Extraneous-Same-Day
2: 1511 775155   0     889 HEIGHTCM       89.06                     Include
3: 1518 775155   0     889 WEIGHTKG       13.10                     Include
4: 1512 775155   0    1071 HEIGHTCM       92.50                     Include
5: 1519 775155   0    1071 WEIGHTKG       14.70                     Include
6: 1513 775155   0    1253 HEIGHTCM       96.20                     Include
longwide(my_cleaned_data, agedays = "aged", param = "type")
```

By default, `longwide()` will only transform records flagged by `cleangrowth()`
for inclusion. To include more categories assigned by `cleangrowth()`, use the
`inclusion_types` option. For example, to include carried forward values along
with included records for the BMI calculation:

```{r, eval = FALSE}
cleaned_data_wide_cf <- longwide(
  long_df = cleaned_data,
  inclusion_types=c("Include", "Exclude-Carried-Forward")
)
```

An additional option, `include_all`, set to `FALSE` by default, will include all
observations for transformation.

See `longwide()` for full details.

With wide data in hand, output taken directly from `longwide()` can have BMI
added with `simple_bmi()`, and then the output can be passed to `ext_bmiz()`, as
shown in the simple example above. Alternatively, you can provide a similarly
formatted data frame directly to `ext_bmiz()`.

## Recoding sex values

Note that `ext_bmiz()` expects the `sex` variable to be coded as a numeric value
of `1` (male) or `2` (female). This difference from the `growthcleanr` [data
preparation specification](quickstart.html#data-preparation-1.) sustains
compatibility with the CDC SAS macro. The `longwide()` function will handle this
conversion given the expected input values of `0` (male) or `1` (female).

If you are using different input data for `ext_bmiz()`, ensure the `sex`
variable in your data set is properly coded. To do so, use `recode_sex()`. For
example, if you have data in the PCORnet CDM format, and want to prepare it for
`ext_bmiz()`:

```{r, eval = FALSE}
recode_sex(
  input_data = cdm_formatted,
  sourcecol = "sex",
  sourcem = "M",
  sourcef = "F",
  targetm = 1L,
  targetf = 2L
)
```

`recode_sex()` can also be used for other purposes, such as recoding values in
preparation for cleaning with `cleangrowth()`, or transforming `growthcleanr`
output to match external specifications.

## Computing BMI percentiles and Z-scores

With data in wide format with BMI, and with sex variables properly coded,
`ext_bmiz()` can be called:

```{r, eval = FALSE}
cleaned_data_bmiz <- ext_bmiz(cleaned_data_bmi)
head(cleaned_data_bmiz)
           id    agey     agem      bmi sex   wt     ht agedays      age     bmi_l    bmi_m      bmi_s         waz    mod_waz        haz    mod_haz     bmiz  mod_bmiz
1: 1000287994 10.0301 120.3612 22.84087   1 40.1 132.50    3661 120.3612 -2.767881 16.64016 0.12002409  1.09852664  0.8417277 -0.9597118 -0.9717678 1.757419 1.5409867
2: 1000287994 11.0466 132.5592 22.84087   1 40.1 132.50    4032 132.5592 -2.589785 17.20379 0.12676206  0.52509455  0.3595874 -1.6331550 -1.6408514 1.584028 1.2827508
3: 1000287994 12.0630 144.7560 18.35674   1 40.1 147.80    4403 144.7560 -2.444694 17.82828 0.13146572 -0.08603109 -0.1155501 -0.2209003 -0.2267003 0.214444 0.1132188
4: 1000287994 13.0795 156.9540 24.12048   1 57.8 154.80    4774 156.9540 -2.325379 18.49734 0.13420981  1.07298540  0.8792766 -0.2418916 -0.2436118 1.475755 1.1619424
5: 1000287994 14.0959 169.1508 25.03608   1 66.6 163.10    5145 169.1508 -2.222102 19.19533 0.13527053  1.24466369  1.0724802 -0.1772860 -0.1713432 1.483245 1.1877586
6: 1000287994  4.9479  59.3748 16.88341   1 19.5 107.47    1806  59.3748 -2.533423 15.43155 0.07546578  0.48691152  0.3858433 -0.2363398 -0.2340109 1.065532 0.8981828
       bmip      p95      p97    bmip95       wp       hp obese sev_obese ext_bmiz ext_bmip    sigma
1: 96.05768 22.14172 23.71268 103.15762 86.40127 16.86001     1         0 1.705525 95.59517 4.668954
2: 94.34063 23.21872 24.93200  98.37264 70.02413  5.12182     0         0 1.584028 94.34063 5.002164
3: 58.48996 24.25081 26.05464  75.69536 46.57209 41.25850     0         0 0.214444 58.48996 5.316539
4: 92.99952 25.21239 27.05537  95.66916 85.83612 40.44321     0         0 1.475755 92.99952 5.612141
5: 93.09953 26.09135 27.92776  95.95547 89.33723 42.96419     0         0 1.483245 93.09953 5.888911
6: 85.66823 17.91161 18.39937  94.25957 68.68395 40.65845     0         0 1.065532 85.66823 2.720945
```

The output columns include:

| variable                  | description                                                                                                                                                                       |
| ------------------------- | --------------------------------------------------------------------------------------------------------------------------------------------------------------------------------- |
| waz, wp                   | LMS Weight-for-sex/age z-score and percentile                                                                                                                                     |
| haz, hp                   | LMS Height-for-sex/age z-score and percentile                                                                                                                                     |
| bmiz, bmip                | LMS BMI-for-sex/age z-score and percentile                                                                                                                                        |
| mod_waz, mod_haz, mod_bmi | Modified z-scores for identifying outliers (see the information in the [CDC SAS growth charts program website](https://www.cdc.gov/nccdphp/dnpao/growthcharts/resources/sas.htm)) |
| bmip95                    | BMI expressed as a percentage of the 95th percentile. A value ≥ 120 is widely used as the cut point for severe obesity.                                                           |
| sigma                     | Scale parameter of the half-normal distribution                                                                                                                                   |
| ext_bmip                  | Extended BMI percentile                                                                                                                                                           |
| ext_bmiz                  | Extended BMI z-score                                                                                                                                                              |

For convenience, these labels are available on the output of `ext_bmiz()`, e.g., when
viewed in RStudio with `View(cleaned_data_bmi)`.

Like `longwide()`, `ext_bmiz()` also includes options for mapping alternate column
names, for age, weight, height, and BMI. The default column names are the same as the
output from `longwide()` for convenience. If you have different column names, specify
the column names without quotes. For example, for a dataset using "heightcm" and
"weightkg" instead of "ht" and "wt":

```{r, eval = FALSE}
my_cleaned_data_bmiz <- ext_bmiz(my_cleaned_data_wide_bmi, ht = heightcm, wt = weightkg)
```

For `ext_bmiz()`, use the most precise age in months available. If an input dataset only
has age in months as integer values, by default `ext_bmiz()` will automatically convert
these to double values and add `0.5` to account for the distribution of actual ages over
the range of days within a month. This is enabled with the option `adjust.integer.age`,
set to `TRUE` by default. Specify `FALSE` to disable.

```{r, eval = FALSE}
my_cleaned_data_bmi <- ext_bmiz(my_cleaned_data_wide, adjust.integer.age = FALSE)
```

`ext_bmiz()` uses reference data provided by the CDC, included in the `growthcleanr`
package as `inst/extdata/CDCref_d.csv`. This file is automatically loaded and used by
default. If you are working with a different reference dataset or developing the
`growthcleanr` package, specify an alternate path to this file with `ref.data.path`, as
for `cleangrowth()`.

## Related tools

The CDC provides a [SAS Program for the 2000 CDC Growth
Charts](https://www.cdc.gov/nccdphp/dnpao/growthcharts/resources/sas.htm) which can
also be used to identify biologically implausible values using a different approach, as
also implemented for `growthcleanr` in the function `ext_bmiz()`, described above.

[GrowthViz](https://github.com/mitre/GrowthViz) provides insights into how
`growthcleanr` assesses data, packaged in a Jupyter notebook. It ships with the
same `syngrowth` synthetic example dataset as `growthcleanr`, with cleaning
results included.
