---
title: "Utilities for computing pediatric BMI percentiles, Z-scores, and related tools"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Utilities}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

The CDC provides a
[SAS macro for computing BMI percentiles and Z-scores](https://www.cdc.gov/nccdphp/dnpao/growthcharts/resources/sas.htm).
The function `ext_bmiz()`, included in `growthcleanr`, provides an equivalent
feature. `ext_bmiz()` calculates the sigma (scale parameter for the half-normal
distribution, extended BMI percentile), extended BMIz, and the CDC LMS z-scores
for weight, height, and BMI. Note that for BMIs ≤ 95th percentile of the CDC
growth charts, the extended values for BMI are equal to the LMS values. The
extended values differ only for children who have a BMI > 95th percentile.

The function assumes a variable ‘sex’

and variables for age in months, weight (kg), height (cm), and BMI (weight/ht2).
Please be careful with age - the units should be months and use the most
accurate information available (e.g., 23.4928 months. The extended BMIz is the
inverse cumulative distribution function (CDF) of the extended BMI percentile.
If the extended percentile is very close to 100, the `qnorm` function in R
produces an infinite value. The occurs only if the extended BMI percentile is >
99.99999999999999. This occurs infrequently, such as a 48-month-old with a BMI >
39, and it is likely that these BMIs represent data entry errors. For these
cases, extended BMIz is set to 8.21, a value that is slightly greater than the
largest value that can be calculated.

The CDC SAS macro was updated in December 2022, according to the findings of
[this NCHS report](https://dx.doi.org/10.15620/cdc:121711). The `ext_bmiz()`
function has been updated to match it as of growthcleanr v2.1.0.

## Converting long growthcleanr data to wide format w/BMI

Because `ext_bmiz()` performs cross-sectional analysis of BMI, observation data
must be in a wide format, i.e. with height and weight information on the same
row. This is distinct from `cleangrowth()`, which performs longitudinal analysis
on all observations for each subject, presented in a long format with one
observation per row. To facilitate use of both functions, `growthcleanr`
includes utility functions to transform data used with `cleangrowth()` for use
with `ext_bmiz()`. They are optimized to move data directly from the output of
`cleangrowth()` into input for `ext_bmiz()`, but have options to support
independent use as well.

Using the `syngrowth` example dataset, to convert the data after it has been
cleaned by `cleangrowth()` for use with `ext_bmiz()`, use `longwide()` and
`simple_bmi()`:

```{r, eval = FALSE}
# Use the built-in utility function to convert the input observations to wide
# format for BMI calculation
cleaned_data_wide <- longwide(cleaned_data)

# Compute simple BMI values (adds column "bmi")
cleaned_data_bmi <- simple_bmi(cleaned_data_wide)

# Compute Z-scores and percentiles
cleaned_data_bmiz <- ext_bmiz(cleaned_data_bmi)
```

Note that this assumes that `cleaned_data` has the same structure as described
in [Quickstart - Data preparation](quickstart.html#data-preparation-1):

```{r, eval = FALSE}
names(cleaned_data)
[1] "id"          "subjid"      "sex"         "agedays"     "param"       "measurement" "gcr_result"
```

The wide dataset `cleaned_data_wide` will include rows with aligned height and
weight measurements drawn from the observations in `cleaned_data` marked by
`cleangrowth()` for inclusion. As such, it will be a shorter dataset (fewer
rows) based on fewer observations.

```{r, eval = FALSE}
dim(cleaned_data)
[1] 85728     7

dim(cleaned_data_wide)
[1] 26701     9

head(cleaned_data_wide)
                                subjid    agey     agem sex   wt wt_id    ht ht_id  agedays
1 002986c5-354d-bb9d-c180-4ce26813ca28 56.0964 673.1568   2 71.7 83331 151.1 83330 20489.22
2 002986c5-354d-bb9d-c180-4ce26813ca28 57.1122 685.3464   2 73.2 83333 151.1 83332 20860.22
3 002986c5-354d-bb9d-c180-4ce26813ca28 58.1279 697.5348   2 74.6 83336 151.1 83335 21231.22
4 002986c5-354d-bb9d-c180-4ce26813ca28 59.1437 709.7244   2 72.8 83338 151.1 83337 21602.22
5 002986c5-354d-bb9d-c180-4ce26813ca28 59.2012 710.4144   2 72.4 83340 151.1 83339 21623.22
6 002986c5-354d-bb9d-c180-4ce26813ca28 60.1594 721.9128   2 69.4 83343 151.1 83342 21973.22
```

In this example, the subject identifiers previously marked as `subjid` are now
in the `id` column; individual identifiers for observations of a single
parameter are not present.

`longwide()` can be called with name mapping parameters if your input set uses
different column names. For example, if `my_cleaned_data` specifies age in days
as `aged` and parameter type as `type`, specify each, with quotes:

```{r, eval = FALSE}
head(my_cleaned_data)
     id subjid sex    aged     type measurement                  gcr_result
1: 1510 775155   0     889 HEIGHTCM       84.90 Exclude-Extraneous-Same-Day
2: 1511 775155   0     889 HEIGHTCM       89.06                     Include
3: 1518 775155   0     889 WEIGHTKG       13.10                     Include
4: 1512 775155   0    1071 HEIGHTCM       92.50                     Include
5: 1519 775155   0    1071 WEIGHTKG       14.70                     Include
6: 1513 775155   0    1253 HEIGHTCM       96.20                     Include
longwide(my_cleaned_data, agedays = "aged", param = "type")
```

By default, `longwide()` will only transform records flagged by `cleangrowth()`
for inclusion. To include more categories assigned by `cleangrowth()`, use the
`inclusion_types` option. For example, to include carried forward values along
with included records for the BMI calculation:

```{r, eval = FALSE}
cleaned_data_wide_cf <- longwide(
  long_df = cleaned_data,
  inclusion_types=c("Include", "Exclude-Carried-Forward")
)
```

An additional option, `include_all`, set to `FALSE` by default, will include all
observations for transformation.

See `longwide()` for full details.

With wide data in hand, output taken directly from `longwide()` can have BMI
added with `simple_bmi()`, and then the output can be passed to `ext_bmiz()`, as
shown in the simple example above. Alternatively, you can provide a similarly
formatted data frame directly to `ext_bmiz()`.

## Recoding sex values

Note that `ext_bmiz()` allows for the `sex` variable to be coded using
[a range of possible values](#computing-bmi-percentiles-and-z-scores), but not
the same `0` and `1` values as `cleangrowth()`. This difference from the
`growthcleanr`
[data preparation specification](quickstart.html#data-preparation-1.) sustains
compatibility with the CDC SAS macro. The `longwide()` function will handle this
conversion from `growthcleanr`'s `0` (male) or `1` (female), but not from other
coded values.

If you are using input data with different value codes for `sex` with
`ext_bmiz()`, use `recode_sex()` to ensure your values are recoded first. For
example, if you have data in the PCORnet CDM format (using `M` and `F`), and
want to prepare it for `ext_bmiz()`:

```{r, eval = FALSE}
recode_sex(
  input_data = cdm_formatted,
  sourcecol = "sex",
  sourcem = "M",
  sourcef = "F",
  targetm = 1L,
  targetf = 2L
)
```

`recode_sex()` can also be used for other purposes, such as recoding values in
preparation for cleaning with `cleangrowth()`, or transforming `growthcleanr`
output to match external specifications.

## Computing BMI percentiles and Z-scores

With data in wide format with BMI, and with `sex` values properly coded (as any
of '1', 'b', 'B', 'Boys', 'm', 'M', 'male', or 'Male' for male subjects and any
of '2', 'g', 'G', 'Girls', 'f', 'F', 'female', or 'Female' for female subjects),
`ext_bmiz()` can be called:

```{r, eval = FALSE}
cleaned_data_bmiz <- ext_bmiz(cleaned_data_bmi)
head(cleaned_data_bmiz)
                                 subjid    agey      age   sex    wt wt_id    ht ht_id agedays      bmi      bmiz
                                 <char>   <num>    <num> <int> <num> <int> <num> <int>   <int>    <num>     <num>
1: 001aa16d-bf0e-a077-3b3d-5ab8b58545ad 10.0233 120.2796     2  35.4    17 141.6    15    3661 17.65537 0.3236612
2: 001aa16d-bf0e-a077-3b3d-5ab8b58545ad 11.0390 132.4680     2  39.2    19 147.9    18    4032 17.92048 0.1734315
3: 001aa16d-bf0e-a077-3b3d-5ab8b58545ad 12.0548 144.6576     2  44.8    21 155.1    20    4403 18.62320 0.1832443
4: 001aa16d-bf0e-a077-3b3d-5ab8b58545ad 12.5914 151.0968     2  47.8    23 158.7    22    4599 18.97903 0.1829183
5: 001aa16d-bf0e-a077-3b3d-5ab8b58545ad 13.0705 156.8460     2  50.5    26 160.8    24    4774 19.53077 0.2586449
6: 001aa16d-bf0e-a077-3b3d-5ab8b58545ad  3.9288  47.1456     2  16.6     2 102.6     1    1435 15.76933 0.3453978
       bmip       waz       wp       haz       hp      p95      p97   bmip95   mod_bmiz   mod_waz   mod_haz
      <num>     <num>    <num>     <num>    <num>    <num>    <num>    <num>      <num>     <num>     <num>
1: 62.69027 0.3498817 63.67862 0.5140553 69.63933 22.96109 24.57902 76.89254 0.18485501 0.2399201 0.5008944
2: 56.88438 0.2311645 59.14065 0.5002225 69.15408 24.13836 25.90525 74.24067 0.09543668 0.1556259 0.4955022
3: 57.26968 0.3237374 62.69316 0.4803298 68.45036 25.26981 27.17179 73.69745 0.10065274 0.2201655 0.4855100
4: 57.25689 0.3812700 64.84986 0.5153244 69.68368 25.83904 27.80781 73.45100 0.10011666 0.2596479 0.5212281
5: 60.20454 0.4440465 67.14955 0.4849566 68.61464 26.32757 28.35405 74.18371 0.14353771 0.3024342 0.4885546
6: 63.51023 0.4369963 66.89430 0.5348018 70.36065 18.02950 18.60078 87.46409 0.24462581 0.3332011 0.5224816
      sigma original_bmip original_bmiz sev_obese obese
      <num>         <num>         <num>     <int> <int>
1: 4.443536      62.69027     0.3236612         0     0
2: 4.797031      56.88438     0.1734315         0     0
3: 5.148292      57.26968     0.1832443         0     0
4: 5.332930      57.25689     0.1829183         0     0
5: 5.497248      60.20454     0.2586449         0     0
6: 2.274792      63.51023     0.3453978         0     0
```

The output columns include:

| variable                     | description                                                                                                                                                                                                                       |
| ---------------------------- | --------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------- |
| bmi                          | BMI                                                                                                                                                                                                                               |
| bmiz, bmip                   | LMS / Extended z-score and percentile                                                                                                                                                                                             |
| waz, wp                      | LMS Weight-for-sex/age z-score and percentile                                                                                                                                                                                     |
| haz, hp                      | LMS Height-for-sex/age z-score and percentile                                                                                                                                                                                     |
| p95, p97                     | 95th and 97th percentile of BMI in growth charts                                                                                                                                                                                  |
| bmip95                       | BMI expressed as a percentage of the 95th percentile. A value ≥ 120 is widely used as the cut point for severe obesity.                                                                                                           |
| mod_bmiz, mod_waz, mod_haz   | Modified BMI-for-age, Weight-for-age, and Height-for-age z-scores for identifying outliers (see the information in the [CDC SAS growth charts program website](https://www.cdc.gov/nccdphp/dnpao/growthcharts/resources/sas.htm)) |
| sigma                        | Scale parameter for half-normal distribution                                                                                                                                                                                      |
| original_bmiz, original_bmip | LMS BMI-for-sex/age z-score and percentile                                                                                                                                                                                        |
| sev_obese                    | BMI >= 120% of 95th percentile (0/1)                                                                                                                                                                                              |
| obese                        | BMI >= 95th percentile (0/1)                                                                                                                                                                                                      |

For convenience, these labels are available on the output of `ext_bmiz()`, e.g.,
when viewed in RStudio with `View(cleaned_data_bmi)`.

Like `longwide()`, `ext_bmiz()` also includes options for mapping alternate
column names, for age, weight, height, and BMI. The default column names are the
same as the output from `longwide()` for convenience. If you have different
column names, specify the column names without quotes. For example, for a
dataset using "heightcm" and "weightkg" instead of "ht" and "wt":

```{r, eval = FALSE}
my_cleaned_data_bmiz <- ext_bmiz(my_cleaned_data_wide_bmi, ht = heightcm, wt = weightkg)
```

For `ext_bmiz()`, use the most precise age in months available. If an input
dataset only has age in months as integer values, by default `ext_bmiz()` will
automatically convert these to double values and add `0.5` to account for the
distribution of actual ages over the range of days within a month. This is
enabled with the option `adjust.integer.age`, set to `TRUE` by default. Specify
`FALSE` to disable.

```{r, eval = FALSE}
my_cleaned_data_bmi <- ext_bmiz(my_cleaned_data_wide, adjust.integer.age = FALSE)
```

`ext_bmiz()` uses reference data provided by the CDC, included in the
`growthcleanr` package as `inst/extdata/CDCref_d.csv`. This file is
automatically loaded and used by default. If you are working with a different
reference dataset or developing the `growthcleanr` package, specify an alternate
path to this file with `ref.data.path`, as for `cleangrowth()`.

## Related tools

The CDC provides a
[SAS Program for the 2000 CDC Growth Charts](https://www.cdc.gov/nccdphp/dnpao/growthcharts/resources/sas.htm)
which can also be used to identify biologically implausible values using a
different approach, as also implemented for `growthcleanr` in the function
`ext_bmiz()`, described above. The SAS program was updated in December, 2022,
and this function has been updated to match it as of growthcleanr v2.1.0.

[GrowthViz](https://github.com/mitre/GrowthViz) provides insights into how
`growthcleanr` assesses data, packaged in a Jupyter notebook. It ships with the
same `syngrowth` synthetic example dataset as `growthcleanr`, with cleaning
results included.
